---
title: "Readme"
output: html_document
---

This file provides the details of run_analysis.r script. I've split this into 7 parts, which do not necessarily follow the exact order that was detailed in out in the instructions.

### Part 1 - Reading the data

The first part of the code reads in the following data files from into data frames. All files can be found in /data directory.

Data set | Data Frame
------------- | -------------
X_test.txt | test_xDF
subject_test.txt | sub_testDF
y_test.txt | test_yDF
X_train.txt | train_xDF
subject_train.txt | sub_trainDF
y_train.txt | train_yDF



### Part 2 - Merging the data

To merge the train and test data frames we use 
```{r eval=FALSE} 
cbind()
```
to join all the 'test' data frames (testJoin) and 'train' data frames (trainJoin). We then use 
```{r eval=FALSE} 
rbind()
```
to merge testJoin and trainJoin data frames in to 'MergedData' table.

### Part 3 - Creating headers  

Next we read in the column labels from 'features.txt' and apply these to the MergedData table.

### Part 4 - Extracting mean() and std() columns
'MeanStd' data frame contains all the mean() and std() [standard deviation] columns from MergedData.  This was executed using the 
```{r eval=FALSE} 
grepl() 
```
command being careful to select patterns where 
```{r eval=FALSE} 
"mean\\("
"std\\("
```
This avoided also selecting columns like "...meanfreq()" which dont have a std() counterpart. I think which "mean" columns you select is a bit subjective anyway. 

With this criterion there should be 68 columns in total that include the activity and subject columns.

Also added is a column containing the activity labels e.g. WALKING, LAYING etc... 

### Part 5 - Tidying the header
Here we apply some text formatting to the MeanStd header as per guidance from Week 4 lectures:

* Force all labels to be lower case.
* Removed all '()' charachters.
* Replaced std with standarddeviation
* Removed '-' character.

The full description of the column labels can be found in CodeBook.txt file.



### Part 6 - Computing the averages and writing to file

We used 
```{r eval=FALSE} 
groupby()
summarise()
```
to compute the average quantaise grouped by subject and activity and wrote the tidied dataframe, 'tidyDF' to a file called 'tidyData.txt'.

This new data frame is called 'Averages'

