---
title: "Readme"
output: html_document
---

This file provides the details of run_analysis.r script. I've split this into 7 parts, which do not necessarily follow the exact order that was detailed in out in the instructions.

### Part 1 - Reading the data

The first part of the code reads in the following data files from  'test' and 'train' into data frames: 

Data set | Data Frame
------------- | -------------
X_test.txt | test_xDF
subject_test.txt | sub_testDF
y_test.txt | test_yDF
X_train.txt | train_xDF
subject_train.txt | sub_trainDF
y_train.txt | train_yDF

### Part 2 - Creating headers  

Next we read in the column labels from 'features.txt' and apply these names to test_xDF and train_yDF. We also create header names for  sub_testDF, test_yDF, sub_trainDF and train_yDF. We will tidy names up later on.

### Part 3 - Merging the data

To merge the train and test data frames we use 
```{r eval=FALSE} 
cbind()
```
to join all the 'test' data frames (testJoin) and 'train' data frames (trainJoin). We then use 
```{r eval=FALSE} 
rbind()
```
to merge testJoin and trainJoin data frames in to 'MergedData'.

### Part 4 - Extracting mean() and std() columns
'MeanStd.df' data frame contains all the mean() and std() [standard deviation] columns from MergedData. 

There should be 66 columns in total.

### Part 5 - Computing the averages 
```{r eval=FALSE} 
colMeans()
```
was used to compute the average column values across the MeanStd.df data frame from Part 4. 

This new data frame is called 'Averages'

### Part 6 - Tidying the header
Here we apply some text formatting to the Averages header as per guidance from Week 4 lectures:

* Checked for duplicate column names (none found)
* Force all labels to be lower case.
* Removed all '()' charachters.
* I chose not to remove '-' characters as I think keeping them in makes the labels easier to read when you want to quickly scan them.

The full description of the column labels can be found in CodeBook.txt file.
